---
title: "RR - Severe Weather Events"
author: "Mike Higgins"
date: "4 July 2016"
output: html_document
---

###Synopsis
This is the synopsis.

### Data Processing
There should be a section titled Data Processing which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the ğšŒğšŠğšŒğš‘ğš = ğšƒğšğš„ğ™´ option for certain code chunks.

Extract the file and make sure to cache this for performance reasons

```{r, cache=TRUE}
sourceFile <- "repdata%2Fdata%2FStormData.csv.bz2"

df <- read.csv(sourceFile)

str(df)

```

Once we have the file loaded into a dataframe, we can go to work preparing and
processing some of the data to make it more usable.
```{r, cache=TRUE}
# Convert date + time columns to a single POSIX date-time column
df$BGN_DATE <- as.POSIXct(df$BGN_DATE, format="%m/%d/%Y %H:%M:%S")
df$BGN_DT <- as.POSIXct(paste(df$BGN_DATE, df$BGN_TIME), format="%Y-%m-%d %H%M")

df$END_DATE <- as.POSIXct(df$END_DATE, format="%m/%d/%Y %H:%M:%S")
df$END_DT <- as.POSIXct(paste(df$END_DATE, df$END_TIME), format="%Y-%m-%d %H%M")

# function to convert damage values to numeric column
conv <- function(x,y) {
  
  x <- as.numeric(x);
  
  if(y == "K") {
    return(x * 1000)
  }
  else if (y == "M") {
    return(x * 1000000)
  }
  else {
    return(x)
  }
  
}

# Create new column of the dollar values
df$PROPDMGVAL <- mapply(conv, df$PROPDMG, df$PROPDMGEXP)

df$CROPDMGVAL <- mapply(conv, df$CROPDMG, df$CROPDMGEXP)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What are we trying to understand?
1. Across the United States, which types of events (as indicated in the ğ™´ğš…ğšƒğšˆğ™¿ğ™´ variable) are most harmful with respect to population health?

- Perhaps we could aggregate deaths and fatalities by EVTYPE and see what the chart 
looks like?


```{r}
propVal <- aggregate(x=df$PROPDMGVAL, by=list(df$EVTYPE), FUN = "sum")
names(propVal) <- c("EVTYPE", "DMGVAL")

cropVal <- aggregate(x=df$CROPDMGVAL, by=list(df$EVTYPE), FUN = "sum")
names(cropVal) <- c("EVTYPE", "DMGVAL")
```


2. Across the United States, which types of events have the greatest economic consequences?
- Perhaps we could aggregate Damage Values by EVTYPE and see what the chart 
looks like?

```{r}
# aggregate number of deaths over various event types
deaths <- aggregate(x=df$FATALITIES, by=list(df$EVTYPE), FUN = "sum")
names(deaths) <- c("EVTYPE", "val")

# aggregate number of injuries over various event types
injuries <- aggregate(x=df$INJURIES, by=list(df$EVTYPE), FUN = "sum")
names(injuries) <- c("EVTYPE", "val")
```
### Results

Should include at least one and no more than 3 charts or figures


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
